Constructing Statistical Deformable Models with Shape Flow

### 2015-07-17 contents/evaluation.tex
Changes for Rebuttal Highlighted

The Cumulative Error Distribution (CED) for this experiment is shown in Figure \ref{fig:svs_ced} while fitting statistics are reported in Table \ref{tab:svs_stats}. Looking at the results it can be clearly seen that using SVS results in an almost constant $\approx$10\% improvement over the simple binary shape representation for most of the CED graph and reported fitting statistics.

@@

\subsection{Correspondence Bases}
\label{exp:basis}
This experiment investigates the importance of applying subspace constraints to the shape flow estimation stage. To this end, we build and compare the performance of two different dense AAMs. The first one (SVS-AAMs) is built from deformation fields obtained by minimizing the variational shape flow energy proposed in section \ref{varshapeflow}, \ie constraining the shape flow with the correspondence bases described in section \ref{sec:trabasis}. The second dense AAM (SVS-AAMs-unconstrained) is built using deformation fields obtained by replacing the correspondence bases by identity bases so that term \ref{eq:lowrank} in the proposed variational shape flow energy becomes: 
\begin{equation}
\beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\bm{v}_n(\bx)\|^2 \ud \bx \label{}
\end{equation}
with the rest of terms (\ref{eq:costfunc} and \ref{eq:TVterm}) remaining the same. Note that this corresponds to a degenerate case where shape flow becomes equivalent to classical frame-to-frame optical flow using TV regularization on the flow $\bm{u}_n$ \cite{Garg:2013hu}.

As before, both dense AAM are build using the 811 training images of LFPW database and tested on the 224 testing images of the same database. The cumulative error distributions graph for this experiment is shown in Figure~\ref{fig:basis_ced} and the fitting statistics reported in Table~\ref{tab:basis_stats}. 
Both Figure~\ref{fig:basis_ced} and Table~\ref{tab:basis_stats} show that the incorporation of the correspondence bases leads to considerably better performant dense AAMs. This large difference in performance fully justifies the incorporation of these bases to our shape flow algorithm.

@@

In the following set of experiments we use the proposed methodology to build dense AAMs of 2 different object classes (faces and ears) and quantitatively assess their respective fitting accuracy when fitted to novel in-the-wild images of the previous objects.

>>>>

The Cumulative Error Distribution (CED) for this experiment is shown in Figure \ref{fig:svs_ced} while fitting statistics are reported in Table \ref{tab:svs_stats}. \textcolor{red}{Looking at the results it can be clearly seen that using SVS results in an almost constant $\approx$10\% improvement over the simple binary shape representation for most of the CED graph and reported fitting statistics.}

@@

\subsection{\textcolor{red}{Correspondence Bases}}
\label{exp:basis}
\textcolor{red}{
This experiment investigates the importance of applying subspace constraints to the shape flow estimation stage. To this end, we build and compare the performance of two different dense AAMs. The first one (SVS-AAMs) is built from deformation fields obtained by minimizing the variational shape flow energy proposed in section \ref{varshapeflow}, \ie constraining the shape flow with the correspondence bases described in section \ref{sec:trabasis}. The second dense AAM (SVS-AAMs-unconstrained) is built using deformation fields obtained by replacing the correspondence bases by identity bases so that term \ref{eq:lowrank} in the proposed variational shape flow energy becomes: 
\begin{equation}
\beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\bm{v}_n(\bx)\|^2 \ud \bx \label{}
\end{equation}
with the rest of terms (\ref{eq:costfunc} and \ref{eq:TVterm}) remaining the same. Note that this corresponds to a degenerate case where shape flow becomes equivalent to classical frame-to-frame optical flow using TV regularization on the flow $\bm{u}_n$ \cite{Garg:2013hu}.}

\textcolor{red}{As before, both dense AAM are build using the 811 training images of LFPW database and tested on the 224 testing images of the same database. The cumulative error distributions graph for this experiment is shown in Figure~\ref{fig:basis_ced} and the fitting statistics reported in Table~\ref{tab:basis_stats}. 
Both Figure~\ref{fig:basis_ced} and Table~\ref{tab:basis_stats} show that the incorporation of the correspondence bases leads to considerably better performant dense AAMs. This large difference in performance fully justifies the incorporation of these bases to our shape flow algorithm.}

@@

\textcolor{red}{In the following set of experiments we use the proposed methodology to build dense AAMs of 2 different object classes (faces and ears) and quantitatively assess their respective fitting accuracy when fitted to novel in-the-wild images of the previous objects.}

### 2015-07-17 contents/implementation.tex
layout fix
Fix eq 6, 7
Changes for Rebuttal Highlighted

\begin{figure*}[t!]
    \centering
        \includegraphics[width=\textwidth]{resources/architecture2}
    \caption{Schematic description of the proposed pipeline. Figure best viewed by zooming in.}
    \label{fig:archi}
\end{figure*}

@@

The proposed pipeline is depicted in Figure~\ref{fig:archi}. It takes as input a set of training images of a particular object class, along with inconsistent point and/or curve line shape annotations.
As a first step, the all shape annotations are consistently represented using multichannel SVS.
The next step is the application of ICP to achieve an initial simple alignment of the SVS images.
The multichannel similarity-aligned SVS images are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all the shapes in the training set. In order for the flow estimation to be accurate and exploit the existent correlation among all training shapes, it utilises a correspondence bases built from sparse correspondences obtained by applying Nonrigid ICP on the annotated data. Finally, the dense correspondences yield by optical flow serve as automatically generated dense landmark annotations and are used to define a novel dense version of Active Appearance Models. The upcoming sections discuss each of the previous stages in further detail.

@@

In order to fully capture the variability among most deformable object shapes annotations, we use a representation based on Support Vector Shapes (SVS) \cite{Nguyen2013}. A SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a) the representation is completely general, \eg it can be applied to sparse landmark points, curves lines or a combination of the two and (b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. Furthermore, this representations is also robust against noise, missing data and outliers \cite{Nguyen2013}.

@@

SVMs with RBF kernel function map any number of data points onto an infinite-dimensional space where positive and negative points are linearly separable, hence the classification boundary on 2D space represents the actual shape of the object. Note that the decision function for SVMs can be mathematically expressed as:
\begin{equation} \label{eq:decisionfunc}
    d(\bm{x})=\sum_i\alpha_i \, k(\bm{x}_i^*,\bm{x})
\end{equation}
where $\bm{x}_i^*$ are the so-called support vectors and \mbox{$k(\bm{x}, \bm{y}) = \exp\left(-\frac{\|\bm{x} -\bm{y}\|^2}{2 \sigma^2}\right)$} is the RBF kernel .

@@

\begin{figure}[b!]
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/landmark}
        \caption{Spare landmarks}
    \end{subfigure}
    \qquad
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/svs}
        \caption{Decision function}
        \label{fig:svs}
    \end{subfigure}
    \caption{Exemplar SVS shape representation. The decision function is trained on the set of sparse landmarks. In (b), brighter colour represents higher probability of a pixels belonging to the original shape.}
    \label{fig:build_svs}
\end{figure}

@@

We consider the problem of dense correspondences across SVS images as a joint estimation of optical flow fields between a reference SVS image and every SVS image of the training dataset. The estimation of such a ``shape flow'' defines for every training SVS image a warping function that registers it with the reference SVS image. To establish the dense correspondences robustly, we are inspired by the idea of subspace constraints in the estimation of multiframe optical flow \cite{Garg:2013hu}.

Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constraints on how points of different shapes are matched to each other. Every pixel of the reference SVS image is matched to its corresponding position at every training SVS image and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all SVS images. To form this vector, the training images are arranged in an arbitrary order. Similarly to the order of training samples when Principal Component Analysis (PCA) is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.

Formally, let $F$ be the number of training SVS images and $n=1,\ldots,F$ be the training image index. Also, let $\boldsymbol{q}_1(n),\ldots,\boldsymbol{q}_R(n):
\{1,\ldots,F\} \rightarrow \R^2$ be the $R$ orthonormal elements of the correspondence basis, where $\boldsymbol{q}_i(n)$ is the displacement vector that matches the points of the reference SVS image with the points of the $n$-th training SVS image, according to the variation described from the $i$-th correspondence basis element. Note that the basis elements $\boldsymbol{q}_i(n)$ are independent from the point location. Note also that the number of basis elements is typically much smaller than the full dimensionality ($2 F$) of correspondence vectors, therefore this basis plays a role of dimensionality reduction.

In addition, let $\Omega \subset \R^2$ be the image domain of the SVS images and $\bx$ denote the point location. We denote the shape flow result as $\bm{u}_n(\bx):\Omega\times \{1,\ldots,F\}
\rightarrow \R^2$,  where $\bm{u}_n(\bx)$ is the displacement vector that matches the point $\bx$ of the reference SVS image with its corresponding location at the $n$-th training SVS image.

Using the constructed correspondence basis, the shape flow can be approximated as:
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx) \,\, ,
\end{equation}
where $\bm{v}_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v}_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v}_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v}_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/2}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/5}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/6}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/7}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.15\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/8}
    \end{subfigure}
    \caption{Exemplar deformation fields obtained using the proposed pipeline. Figure best viewed by zooming in.}
    \label{fig:deformationfield}
\end{figure}

@@

Using the SVS functions built from training data, one of the main steps of our pipeline is to estimate dense correspondences between all these functions. We propose to perform this estimation simultaneously for all training data, by using the correspondence basis. This procedure, which we call \emph{shape flow}, builds upon robust methods for multiframe optical flow estimation \cite{Garg:2013hu}. However, optical flow estimation typically works based on the assumptions of brightness or colour constancy and motion smoothness, whereas in our setting the input training data correspond to shapes. For this reason, we propose to modify the formulation of \cite{Garg:2013hu} by using the correspondence basis that we introduced in conjunction with the SVS representation of shapes.

@@

To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply Non-rigid Iterative Closest Point (NICP)\cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. To make this correspondence vectors estimation dense (i.e.~for every pixel of the reference SVS image), we apply Thin-Plate Splines (TPS)\cite{Bookstein1989} interpolation. Finally, the correspondence basis is found by applying PCA on the result of TPS interpolation and keeping only the first $R$ principal components.

@@

Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training SVS images $\bm{d}_n(\bx)$ by minimizing the following energy:
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^F \|\bm{d}_n(\bx+\bm{u}_n(x))-\bm{d}_0(x)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v}_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v}_i(\bx)$ that are spatially regularised and enforce a low-rank prior. We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v}_i(\bx)$. The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

The \textbf{first term} of the above energy \eqref{eq:costfunc} is a data attachment term
that uses the robust $\Lone$-norm.  It is based on the assumption that the values of the reference SVS image $\bm{d}_0(\bx)$ at every pixel $\bx$ are preserved at its corresponding locations on all training SVS images $\bm{d}_n(\bx)$. The use of an $\Lone$-norm improves the robustness of the method since it allows deviations from this assumption, which might occur in practice.
The \textbf{second term} of the energy \eqref{eq:lowrank} penalizes the difference between the two sets  of shape flows and acts as a coupling term between them.
The \textbf{third term} of the energy \eqref{eq:TVterm} corresponds to the spatial Total Variation regularization \cite{rudin92} of
the correspondence basis coefficients $\bm{v}_i(\bx)$.
This term penalizes spatial oscillations of each coefficient caused by distortions of the SVS images but not strong discontinuities that are desirable in the borders of different object regions. In addition, this term allows to fill in information into regions where the shape information in the SVS images is missing, due to e.g.~regions with no annotations.

We implement the minimization of the energy $E_{sf}$ by using the optimization algorithm described in \cite{Garg:2013hu}. In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with the shape flow result of the TPS interpolation described in Sec.~\ref{sec:trabasis}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$\bm{v}_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
$\bm{v}_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.

Figure~\ref{fig:deformationfield} shows some examples of deformation fields derived from the estimated shape flow computed by the aforementioned method. These results correspond to exemplar training shapes in the case of a face dataset.

@@

Note that, both shape and texture models are directly related through the existent one-to-one correspondence between dense landmarks and pixels on the reference
frame. Therefore, in clear contrast to classic AAMs formulations \cite{Cootes2001, Matthews2004}, in our dense formulation no motion model (piece-wise affine, thin-plates splines \cite{Bookstein1989}) is required to relate the previous shape and texture models . As noted by other authors \cite{Amberg2009, Tzimiropoulos2014}, this has the advantage that all efficient compositional gradient descent algorithms for fitting AAMs \cite{Papandreou2008, Matthews2004, Amberg2009, Tzimiropoulos2013, Alabort2014} become exact under our formulation.

>>>>

\begin{figure*}[t!]
    \centering
        \includegraphics[width=\textwidth]{resources/architecture3}
    \caption{\textcolor{red}{Schematic description of the proposed pipeline. Figure best viewed by zooming in.}}
    \label{fig:archi}
\end{figure*}

@@

\textcolor{red}{The proposed pipeline is depicted in Figure~\ref{fig:archi}. It takes as input a set of training images of a particular object class, along with inconsistent point and/or curve line shape annotations.
As a first step, the all shape annotations are consistently represented using multichannel SVS.
The next step is the application of ICP to achieve an initial simple alignment of the SVS images.
The multichannel similarity-aligned SVS images are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all the shapes in the training set. In order for the flow estimation to be accurate and exploit the existent correlation among all training shapes, it utilises a correspondence bases built from sparse correspondences obtained by applying Nonrigid ICP on the annotated data. Finally, the dense correspondences yield by optical flow serve as automatically generated dense landmark annotations and are used to define a novel dense version of Active Appearance Models. The upcoming sections discuss each of the previous stages in further detail.}

@@

In order to fully capture the variability among most deformable object shapes annotations, we use a representation based on Support Vector Shapes (SVS) \cite{Nguyen2013}. A SVS is a decision function trained on shapes using Support Vector Machines (SVMs) with Radial Basis Function (RBF) kernels. In this way, a shape is represented as a classifier function, which has several advantages: (a) the representation is completely general, \eg it can be applied to sparse landmark points, curves lines or a combination of the two and (b) it fuses inconsistent landmarks into consistent and directly comparable decision functions. \textcolor{red}{Furthermore, this representations is also robust against noise, missing data and outliers \cite{Nguyen2013}.}

@@

\textcolor{red}{SVMs with RBF kernel function map any number of data points onto an infinite-dimensional space where positive and negative points are linearly separable, hence the classification boundary on 2D space represents the actual shape of the object. Note that the decision function for SVMs can be mathematically expressed as:
\begin{equation} \label{eq:decisionfunc}
    d(\bm{x})=\sum_i\alpha_i \, k(\bm{x}_i^*,\bm{x})
\end{equation}
where $\bm{x}_i^*$ are the so-called support vectors and \mbox{$k(\bm{x}, \bm{y}) = \exp\left(-\frac{\|\bm{x} -\bm{y}\|^2}{2 \sigma^2}\right)$} is the RBF kernel .}

@@

% \begin{figure}[b!]
%     \centering
%     \begin{subfigure}[b]{0.2\textwidth}
%             \includegraphics[width=\textwidth]{resources/landmark}
%         \caption{Spare landmarks}
%     \end{subfigure}
%     \qquad
%     \begin{subfigure}[b]{0.2\textwidth}
%             \includegraphics[width=\textwidth]{resources/svs}
%         \caption{Decision function}
%         \label{fig:svs}
%     \end{subfigure}
%     \caption{Exemplar SVS shape representation. The decision function is trained on the set of sparse landmarks. In (b), brighter colour represents higher probability of a pixels belonging to the original shape.}
%     \label{fig:build_svs}
% \end{figure}

\begin{figure}[b!]
    \centering
    \includegraphics[width=0.4\textwidth]{resources/svs2}
    \caption{\textcolor{red}{Exemplar SVS shape representation. A similar SVS image is obtained from different annotations of the same shape.}}
    \label{fig:build_svs}
\end{figure}

@@

\textcolor{red}{We consider the problem of dense correspondences across SVS images as a joint estimation of optical flow fields between a reference SVS image and every SVS image of the training dataset. The estimation of such a ``shape flow'' defines for every training SVS image a warping function that registers it with the reference SVS image. To establish the dense correspondences robustly, we are inspired by the idea of subspace constraints in the estimation of multiframe optical flow \cite{Garg:2013hu}.}

\textcolor{red}{Instead of the \emph{motion} basis used in multiframe optical flow formulation of \cite{Garg:2013hu}, we build a \emph{correspondence} basis that introduces constraints on how points of different shapes are matched to each other. Every pixel of the reference SVS image is matched to its corresponding position at every training SVS image and in this way defines a \emph{correspondence vector}. This vector consists of the 2D locations of the specific point in all SVS images. To form this vector, the training images are arranged in an arbitrary order. Similarly to the order of training samples when Principal Component Analysis (PCA) is applied, this order does not affect the result of our method and any re-ordering would produce exactly the same results.}

\textcolor{red}{Formally, let $F$ be the number of training SVS images and $n=1,\ldots,F$ be the training image index. Also, let $\boldsymbol{q}_1(n),\ldots,\boldsymbol{q}_R(n):
\{1,\ldots,F\} \rightarrow \R^2$ be the $R$ orthonormal elements of the correspondence basis, where $\boldsymbol{q}_i(n)$ is the displacement vector that matches the points of the reference SVS image with the points of the $n$-th training SVS image, according to the variation described from the $i$-th correspondence basis element. Note that the basis elements $\boldsymbol{q}_i(n)$ are independent from the point location. Note also that the number of basis elements is typically much smaller than the full dimensionality ($2 F$) of correspondence vectors, therefore this basis plays a role of dimensionality reduction.}

\textcolor{red}{In addition, let $\Omega \subset \R^2$ be the image domain of the SVS images and $\bx$ denote the point location. We denote the shape flow result as $\bm{u}_n(\bx):\Omega\times \{1,\ldots,F\}
\rightarrow \R^2$,  where $\bm{u}_n(\bx)$ is the displacement vector that matches the point $\bx$ of the reference SVS image with its corresponding location at the $n$-th training SVS image.}

\textcolor{red}{Using the constructed correspondence basis, the shape flow can be approximated as:
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx) \,\, ,
\end{equation}
where $\bm{v}_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v}_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v}_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v}_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.}

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/2}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/5}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/6}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/7}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Flows/8}
    \end{subfigure}
    \caption{\textcolor{red}{Exemplar deformation fields obtained using the proposed pipeline. Figure best viewed by zooming in.}}
    \label{fig:deformationfield}
\end{figure}

@@

\textcolor{red}{Using the SVS functions built from training data, one of the main steps of our pipeline is to estimate dense correspondences between all these functions. We propose to perform this estimation simultaneously for all training data, by using the correspondence basis. This procedure, which we call \emph{shape flow}, builds upon robust methods for multiframe optical flow estimation \cite{Garg:2013hu}. However, optical flow estimation typically works based on the assumptions of brightness or colour constancy and motion smoothness, whereas in our setting the input training data correspond to shapes. For this reason, we propose to modify the formulation of \cite{Garg:2013hu} by using the correspondence basis that we introduced in conjunction with the SVS representation of shapes.}

@@

\textcolor{red}{To effectively build the correspondence basis, we first transform the original annotations to sparse point clouds. Then, we apply Non-rigid Iterative Closest Point (NICP)\cite{Amber2007} between the point cloud of annotations in the reference shape and the one of every shape of the training set. NICP iteratively deforms the cloud of points of every shape to match the points of the reference shape. This yields an initial estimation of the correspondence vectors on the sparse locations of annotated landmarks on the reference shape. To make this correspondence vectors estimation dense (i.e.~for every pixel of the reference SVS image), we apply Thin-Plate Splines (TPS)\cite{Bookstein1989} interpolation. Finally, the correspondence basis is found by applying PCA on the result of TPS interpolation and keeping only the first $R$ principal components.}

@@

\textcolor{red}{Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training SVS images $\bm{d}_n(\bx)$ by minimizing the following energy:
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^F \|\bm{d}_n(\bx+\bm{u}_n(x))-\bm{d}_0(x)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^F\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v}_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v}_i(\bx)$ that are spatially regularised and enforce a low-rank prior. We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v}_i(\bx)$. The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.}

\textcolor{red}{The \textbf{first term} of the above energy \eqref{eq:costfunc} is a data attachment term
that uses the robust $\Lone$-norm.  It is based on the assumption that the values of the reference SVS image $\bm{d}_0(\bx)$ at every pixel $\bx$ are preserved at its corresponding locations on all training SVS images $\bm{d}_n(\bx)$. The use of an $\Lone$-norm improves the robustness of the method since it allows deviations from this assumption, which might occur in practice.
The \textbf{second term} of the energy \eqref{eq:lowrank} penalizes the difference between the two sets  of shape flows and acts as a coupling term between them.
The \textbf{third term} of the energy \eqref{eq:TVterm} corresponds to the spatial Total Variation regularization \cite{rudin92} of
the correspondence basis coefficients $\bm{v}_i(\bx)$.
This term penalizes spatial oscillations of each coefficient caused by distortions of the SVS images but not strong discontinuities that are desirable in the borders of different object regions. In addition, this term allows to fill in information into regions where the shape information in the SVS images is missing, due to e.g.~regions with no annotations.}

\textcolor{red}{We implement the minimization of the energy $E_{sf}$ by using the optimization algorithm described in \cite{Garg:2013hu}. In every coarse-to-fine and warping iteration, we use an initialization that comes from the previous iteration and for the very beginning, we initialize with the shape flow result of the TPS interpolation described in Sec.~\ref{sec:trabasis}. We approximate the data term \eqref{eq:costfunc} by linearizing the SVS images around the initialization. After that, the energy becomes convex and we optimize it using alternating optimization w.r.t.~$\bm{v}_i(\bx)$ and $\bm{u}_n(\bx)$. The minimization w.r.t.~
$\bm{v}_i(\bx)$ is decoupled for every coefficient $i$ and corresponds to Rudin-Osher-Fatemi Total Variation denoising \cite{rudin92}, which we solve efficiently by applying the first order primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}. The minimization w.r.t.~
$\bm{u}_n(x)$ is decoupled for every pixel $\bx$ and every shape index $i$. This minimization is also implemented by applying the efficient primal-dual algorithm of \cite{Chambolle:Pock:JMIV2011}.}

\textcolor{red}{Figure~\ref{fig:deformationfield} shows some examples of deformation fields derived from the estimated shape flow computed by the aforementioned method. These results correspond to exemplar training shapes in the case of a face dataset.}

@@

Note that, both shape and texture models are directly related through the existent one-to-one correspondence between dense landmarks and pixels on the reference
frame. \textcolor{red}{Therefore, in clear contrast to classic AAMs formulations \cite{Cootes2001, Matthews2004}, in our dense formulation no motion model (piece-wise affine, thin-plates splines \cite{Bookstein1989}) is required to relate the previous shape and texture models . As noted by other authors \cite{Amberg2009, Tzimiropoulos2014}, this has the advantage that all efficient compositional gradient descent algorithms for fitting AAMs \cite{Papandreou2008, Matthews2004, Amberg2009, Tzimiropoulos2013, Alabort2014} become exact under our formulation.}

### 2015-07-17 contents/introduction.tex
Changes for Rebuttal Highlighted

To illustrate how time consuming careful annotation of the human ear is we lay down our own experience. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore,  training and fatigue largely influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the tasks of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, jiang2009learning, liu2009simultaneous, Zhang2012}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed the limited number of methods that have been applied to the task cannot be directly applied to arbitrary collection of images collected "in-the-wild". On the other hand, the method of \cite{antonakos2014automatic}, which can deal with "in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.

>>>>

To illustrate how time consuming careful annotation of the human ear is we lay down our own experience. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore,  training and fatigue largely influence the annotation accuracy. \textcolor{red}{Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the tasks of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, jiang2009learning, liu2009simultaneous, Zhang2012}}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed the limited number of methods that have been applied to the task cannot be directly applied to arbitrary collection of images collected "in-the-wild". On the other hand, the method of \cite{antonakos2014automatic}, which can deal with "in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.
